Uncertainty in the evolution of social learning

### 2022-05-05 meanNetPayoffs.tex
full outline of analysis, now filling it out
started explaining net payoffs

\begin{figure}
  \caption{Dependence of average net payoff ($\langle \pi \rangle$) at the
    end of last time step of the simulation on environmental
    variability $u$ for several values of low payoff $\pilow$, number of
    behaviors $B$, and lifespan $L$. Average over 1000 trials.} 
  \vspace{3em}
  % $\langle s \rangle$ decreases monotonically for $\pisub{low} = 0.1, 0.5$,
% but noise dominates the results for $\pisub{low}=0.8$. When $\pisub{low} = 0.5$, 
% the sigmoid is not as sharp as for $\pisub{low}=0.1$, meaning that social learning
% is reduced for smaller $u$ and increased for larger $u$.}
  \label{fig:results}
  \centering
  \begin{subfigure}[t]{0.08\textwidth}
    \vspace{-2em}
    \hspace{-4em}
    % \centering
    \begin{tikzpicture}
      \draw[->,thick] (-.1,0)--(1,0) node[above]
        {$B$};
        % {Number of behaviors $B$};
      \draw[->,thick] (0,.1)--(0,-1) node[below]
        {$\pilow$};
        % {Low payoff frequency $\pilow$};
    \end{tikzpicture}

>>>>

\begin{figure}
  \caption{Dependence of average net payoff ($\langle \pi \rangle$) at the
    end of last time step of the simulation on environmental
    variability $u$ for several values of low payoff $\pilow$, number of
    behaviors $B$, and lifespan $L$. Dashed lines mark expected 
    payoff from individual learning $\langle \pi_I \rangle$ calculated
    for each parameter setting (not sensitive to $u$). 
    Note that when the maximum $L$ is 
    8, the y-axis runs from 0 to 8, and when the maximum $L$ is 20, the y-axis
    runs from 0 to 20. Average over 1000 trials.} 
  \label{fig:meanNetPayoffs}
  \vspace{3em}
  % $\langle s \rangle$ decreases monotonically for $\pisub{low} = 0.1, 0.5$,
% but noise dominates the results for $\pisub{low}=0.8$. When $\pisub{low} = 0.5$, 
% the sigmoid is not as sharp as for $\pisub{low}=0.1$, meaning that social learning
% is reduced for smaller $u$ and increased for larger $u$.}
  \centering
  \addtocounter{figure}{-1}  % Hack to get figure numbering correct with subfigure.
  \begin{subfigure}[t]{0.08\textwidth}
    \vspace{-2em}
    \hspace{-4em}
    % \centering
    \begin{tikzpicture}
      \draw[->,thick] (-.1,0)--(1,0) node[above]
        {$B$};
        % {Number of behaviors $B$};
      \draw[->,thick] (0,.1)--(0,-1) node[below]
        {$\pilow$};
        % {Low payoff frequency $\pilow$};
    \end{tikzpicture}

### 2022-05-05 ms.tex
started explaining net payoffs

For social learning to evolve, it must improve upon individual learning. 
We found that different parameter settings change how advantageous social learning
is over and above individual learning. To analyze the effect of uncertainty on
payoffs, it is useful to identify some constant reference values which we will
use to evaluate the benefit of social learning. First, the optimal expected payoff
is the optimal behavior payoff times the lifespan, which we can denote
$\langle \pi^*(L) \rangle = L\cdot\pihigh$. The expected base payoff by chance is 
$\langle \pi_{\text{base}} \rangle = L \left[ (B - 1) \pilow + \pihigh \right]$, which
an agent would receive on average from performing purely random behaviors. 
Agents learn adaptively over their lifespan (via the softmax algorithm in 
our model), whether or not they are social learners. We calculated this
value for reference for each triplet of $\pilow$, $B$, and $L$ 
\mt{calculate this and add to plot as a guide}. 
Social learning evolves when the expected payoff of
social learners exceeds the expected payoff of individual learners.

\mt{Use above framework to analyze results after plot guides added as
promised}

>>>>

For social learning to evolve, it must improve upon individual learning. 
We found that different parameter settings indeed change how advantageous social learning
is compared to individual learning (Figure~\ref{fig:meanNetPayoffs}). 
To analyze the effect of uncertainty on
payoffs, it is useful to identify some constant reference values which we will
use to evaluate the benefit of social learning. First, the optimal expected payoff
is the optimal behavior payoff times the lifespan, which we can denote
$\langle \pi^*(L) \rangle = L\cdot\pihigh$. The expected base payoff by chance is 
$\langle \pi_{\text{base}} \rangle = L \left[ (B - 1) \pilow + \pihigh \right]$, which
an agent would receive on average from performing purely random behaviors. 
Most importantly for understanding the benefit and evolution of social learning,
we calculated by simulation the expected payoff to individual learners for
each triplet ($\pilow$, $B$, $L$), which we denote $\langle \pi_I \rangle$
and plot for reference in Figure~\ref{fig:meanNetPayoffs}. Note $\langle \pi_I
\rangle$ is insensitive to $u$, which only affects the benefit of social
learning.

We observe across parameter settings that social learning outperforms
individual learning when $u$ is small. In fact, when $u=0.0$, social learning 
enables each new generation to know the optimal behavior and achieve the maximum
expected payoff for their lifespan, $\langle \pi^*(L) \rangle$. 
However, in some cases, even when $u$ is small, 
the returns on social learning are truly marginal due to 
relatively high, uniform expected payoffs whether or not agents identify the optimal
behavior, especially when $\pilow = 0.8$. In intermediate cases, examining the decay
of the benefit of social learning informs our understanding of social learning
evolution analyzed above.  As our analysis of social learning evolution suggested,
social learning is only beneficial up to a point. When $\pilow=0.1$, increasing $B$
led both to larger differences between
